---
title: "Chapter 11 number 39"
author: "Karla Palos"
output: github_document
---

Questions: 
test- control 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


An experiment was done to test a method for reducing faults on telephone lines (Welch 1987). Fourteen matched pairs of areas were used. The following table shows the fault rates for the control areas and for the test areas:
```{r}
library("dplyr")
test <-c(676,206,230,256,280,433,337,466,497,512,794,428,452,512)
control <-c(88,570,605,617,653,2913,924,286,1098,982,2346,321,615,519)


differences<- test-control
abs_diff <-abs(differences)
rank <- rank(abs_diff)
signed_rank <- rank*(differences/abs_diff)

df <- data.frame(test, control,differences, abs_diff,rank, signed_rank)
df


#a. Plot the differences versus the control rate and summarize what you see.

plotdiff <- qqplot(control, differences)
plotdiff

#b. Calculate the mean difference,its standard deviation,and a confidence interval.


control_mean<- sum(control)/14
test_mean<- sum(test)/14
differences_mean <- sum(df$differences)/14
#vars
VarC<-var(control/14)
VarT<-var(test/14)
VarDiff <-var(df$differences)

#SD
SDC<- sqrt(VarC)
SDT<-sqrt(VarT)
SDd <- sqrt(VarDiff)

#W_pos
W_pos <- df%>%filter(signed_rank>0)%>%summarise(sum(signed_rank))

#values <-c(control_sum, test_sum, differences_sum, Varx, SDx, Vary, SDy, Vard, SDd, W_pos)
#names<-c("control_mean", "test_mean", "differences_mean", "Varx", "SDx", "Vary", "SDy", "Vard", "SDd", "W_pos")
#name_value<- data.frame(names, values)

```

c. Calculate the median difference and a confidence interval and compare to the
previous result.

```{r}
```
d. Do you think it is more appropriate to use a t test or a non parametric method to
test whether the apparent difference between test and control could be due to chance? Why? Carry out both tests and compare.
```{r}
```