---
title: "Question 46"
author: "Karla Palos"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data of this exercise were gathered as part of a study to estimate the population size of the bowhead whale (Raftery and Zeh 1993). The statistical procedures for estimating the population size along with an assessment of the variability of the estimate were quite involved, and this problem deals with only one aspect of the problem—a study of the distribution of whale swimming speeds. Pairs of sightings and corresponding locations that could be reliably attributed to the same whale were collected, thus providing an estimate of velocity for each whale. Thevelocities,v1,v2,...,v210 (km/h),were converted into times t1,t2,...,t210 to swim 1 km—ti = 1/vi . The distribution of the tiwas then fit by a gamma distribution. The times are contained in the file whales.


```{r}

before <- c(24.6,17,16,10.4,8.2,7.9,8.2,7.9,5.8,5.4,5.1,4.7)
after <- c(10.1,5.7,5.6,3.4,6.5,0.7,6.5,0.7,6.1,4.7,2.0,2.9)
diff=before-after
abs_diff=abs(diff)
rank <- rank(abs_diff)
signed_rank <- rank*(diff/abs_diff)
df <- data.frame(before,after,diff, abs_diff,rank, signed_rank)

```
**a. Make a histogram of the 210 values of t *i* . Does it appear that a gamma distribution would be a plausible model to fit?**
```{r}
path <- file.path("~","Desktop","whales.csv")
myData <- read.csv(file = path, header = TRUE, sep = )
library(ggplot2)
colnames(myData) <- "V1"

myHistogram <- ggplot(myData)+geom_histogram(aes(x=V1), binwidth=.5)+labs(title="Time Whales Take To Swim 1 km") + labs(x="Time (Hours)" )
myHistogram
```

**b. Fit the parameters of the gamma distribution by the method of moments.**
```{r}
alphaMom <-mean(myData$V1)^2/var(myData$V1)
betaMom<- mean(myData$V1)/var(myData$V1)

ggplot(myData) + geom_histogram(aes(x=V1, y=..density..), binwidth=.5, col="white") + stat_function(fun=function(x)+dgamma(x, shape=alphaMom, rate=betaMom), col="white") +labs(title="Time Whales Take To Swim 1 km (MOM)") + labs(x="Time (Hours)" )+ theme(text=element_text(size=10))
```

**c. Fit the parameters of the gamma distribution by maximum likelihood. How do these values compare to those found before?**
```{r}
aMle<-1 
bMle<- aMle/mean(myData$V1)

library(MASS)
estimatedValues<- fitdistr(myData$V1, "gamma", start=list(shape=aMle, rate=bMle))$estimate
alphaMle<-estimatedValues[[1]]
betaMle<-estimatedValues[[2]]
alphaMle
betaMle

ggplot(myData) + geom_histogram(aes(x=V1, y=..density..), binwidth=.5, col="white") + stat_function(fun=function(x)+dgamma(x, shape=alphaMle, rate=betaMle), col="red") +labs(title="Time Whales Take To Swim 1 km (MLE)") + labs(x="Time (Hours)" )+ theme(text=element_text(size=10))
```

**d. Plot the two gamma densities on top of the histogram. Do the fits look reasonable?**
```{r}
ggplot(myData) + geom_histogram(aes(x=V1, y=..density..), binwidth=.5, col="white") + stat_function(fun=function(x)+dgamma(x, shape=alphaMle, rate=betaMle), col="red") +  stat_function(fun=function(x)+dgamma(x, shape=alphaMom, rate=betaMom), col="white") +labs(title="Time Whales Take To Swim 1 km") + labs(x="Time (Hours)" )+ theme(text=element_text(size=10))
```

**e. Estimate the sampling distributions and the standard errors of the parameters fit by the method of moments by using the bootstrap.**
```{r warning=FALSE}
gammaBootstrapMOM <- function(x) {
  alphaMom <- mean(x)^2/var(x) #.799
  betaMom <- mean(x)/var(x) #1.318
  
  bootstrap <- sapply(1:1000, function(i) {
    #samples <- sample(x, length(x), replace=TRUE)
    samples <-rgamma(x, shape=alphaMom, rate = betaMom)

    alphaSample <- mean(samples)^2/var(samples)
    betaSample <- mean(samples)/var(samples)
    return(c(alphaSample, betaSample))
  })
 
  estimatesd <-sapply(1:2, function(i){
    simulatedsd<- sd(bootstrap[i,])
    return(simulatedsd)
  })
  
  alpha<- sapply(1, function(i){
    a<- mean(bootstrap[i,])
    return(a)
  })
  
  beta<- sapply(2, function(i){
    b<- mean(bootstrap[i,])
    return(b)
  })
  
  return(c(estimatesd,alpha,beta))
}

MOMBootstrapEstimator <- gammaBootstrapMOM(myData$V1)
names(MOMBootstrapEstimator)<-c("SE Alpha","SE Beta","Sampling Distribution Alpha Estimate","Sampling Distribution Beta Estimate")
MOMBootstrapEstimator 
```

